---
title: "Intro to basics"
author: "Data Camp MOS "
date: "`r Sys.time()`"
output: 
  html_document: 
    highlight: textmate
    keep_md: yes
    theme: cerulean
    toc: yes
    toc_float: yes
---

## Arithmetic with R

In its most basic form, R can be used as a simple calculator. Consider the following arithmetic operators:

+ Addition: +
+ Subtraction: -
+ Multiplication: *
+ Division: /
+ Exponentiation: ^
+ Modulo: %%

The last two might need some explaining:

The ^ operator raises the number to its left to the power of the number to its right: for example 3^2 is 9.

The modulo returns the remainder of the division of the number to the left by the number on its right, for example 5 modulo 3 or 5 %% 3 is 2.

With this knowledge, follow the instructions to complete the exercise.

Instructions

Type 2^5 in the editor to calculate 2 to the power 5.
Type 28 %% 6 to calculate 28 modulo 6.

Submit the answer and have a look at the R output in the console.
Note how the # symbol is used to add comments on the R code.

```{r}
# An addition
5+5 

# A subtraction
5-5 

# A multiplication
3*5

 # A division
(5+5)/2 

# Exponentiation
2^5
# Add your code here!

# Modulo
17%%4
# Add your code here!
2^5
28 %% 6
```

## Variable assignment

A basic concept in (statistical) programming is called a **variable**.

A variable allows you to store a value (e.g. 4) or an object (e.g. a function description) in R. You can then later use this variable's name to easily access the value or the object that is stored within this variable.

You can assign a value 4 to a variable my_var with the command

```
my_var <- 4
```

Instructions

Over to you: complete the code in the editor such that it assigns the value 42 to the variable x in the editor. Submit the answer. Notice that when you ask R to print x, the value 42 appears.

```{r}
# Assign the value 42 to x
x = 42

# Print out the value of the variable x
x

```

## Variable assignment (2)

Suppose you have a fruit basket with five apples. As a data analyst in training, you want to store the number of apples in a variable with the name my_apples.

Instructions

Type the following code in the editor:

```
my_apples <- 5
```
This will assign the value 5 to my_apples.

Type: `my_apples` below the second comment. This will print out the value of my_apples.

```{r}
# Assign the value 5 to the variable my_apples
my_apples <- 5

# Print out the value of the variable my_apples
my_apples 

```

## Variable assignment (3)

Every tasty fruit basket needs oranges, so you decide to add six oranges. As a data analyst, your reflex is to immediately create the variable my_oranges and assign the value 6 to it. Next, you want to calculate how many pieces of fruit you have in total. Since you have given meaningful names to these values, you can now code this in a clear way:

```
my_apples + my_oranges
```
Instructions

Assign to my_oranges the value 6.

Add the variables my_apples and my_oranges and have R simply print the result.

Assign the result of adding my_apples and my_oranges to a new variable my_fruit.

```{r}
# Assign a value to the variables my_apples and my_oranges
my_apples <- 5
my_oranges <- 6

# Add these two variables together
my_apples + my_oranges

# Create the variable my_fruit
my_fruit <- my_apples + my_oranges
```
## Apples and oranges

Common knowledge tells you not to add apples and oranges. But hey, that is what you just did, no :-)? The my_apples and my_oranges variables both contained a number in the previous exercise. The + operator works with numeric variables in R. If you really tried to add "apples" and "oranges", and assigned a text value to the variable my_oranges (see the editor), you would be trying to assign the addition of a numeric and a character variable to the variable my_fruit. This is not possible.

Instructions

Submit the answer and read the error message. Make sure to understand why this did not work.

Adjust the code so that R knows you have 6 oranges and thus a fruit basket with 11 pieces of fruit.

```{r}
# Assign a value to the variable my_apples
my_apples <- 5 

# Fix the assignment of my_oranges
my_oranges <- 6 

# Create the variable my_fruit and print it out
my_fruit <- my_apples + my_oranges 
my_fruit
```
## Basic data types in R
R works with numerous data types. Some of the most basic types to get started are:

+ Decimal values like 4.5 are called **numerics**.
+ Whole numbers like 4 are called **integers**. Integers are also numerics.
+ Boolean values (TRUE or FALSE) are called **logical**.
+ Text (or string) values are called **characters**.

Note how the quotation marks in the editor indicate that "some text" is a string.

Instructions

Change the value of the:

my_numeric variable to 42.

my_character variable to "universe". Note that the quotation marks indicate that "universe" is a character.

my_logical variable to FALSE.

Note that R is case sensitive!

```{r}
# Change my_numeric to be 42
my_numeric <- 42

# Change my_character to be "universe"
my_character <- "universe"

# Change my_logical to be FALSE
my_logical <- FALSE
```

## What's that data type?

Do you remember that when you added 5 + "six", you got an error due to a mismatch in data types? You can avoid such embarrassing situations by checking the data type of a variable beforehand. You can do this with the `class()` function, as the code in the editor shows.

Instructions

Complete the code in the editor and also print out the classes of my_character and my_logical.

```{r}
# Declare variables of different types
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE 

# Check class of my_numeric
class(my_numeric)

# Check class of my_character
class(my_character)

# Check class of my_logical
class(my_logical)
```

