---
title: "Introduction to R"
author: "Data Camp MOS "
date: "`r Sys.time()`"
output: 
  html_document: 
    highlight: textmate
    theme: cerulean
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Matrices

##What's a matrix?

In R, a matrix is a collection of elements of the same data type (numeric, character, or logical) arranged into a fixed number of rows and columns. Since you are only working with rows and columns, a matrix is called two-dimensional.

You can construct a matrix in R with the 
`r matrix()` function. Consider the following example:

```
matrix(1:9, byrow = TRUE, nrow = 3)

```
In the `r matrix()` function: The first argument is the collection of elements that R will arrange into the rows and columns of the matrix. Here, we use 1:9 which is a shortcut for `r c(1, 2, 3, 4, 5, 6, 7, 8, 9)`.

The argument byrow indicates that the matrix is filled by the rows. If we want the matrix to be filled by the columns, we just place byrow = FALSE.

The third argument nrow indicates that the matrix should have three 
rows.

Instructions

Construct a matrix with 3 rows containing the numbers 1 up to 9, filled row-wise.

```{r}
# Construct a matrix with 3 rows that contain the numbers 1 up to 9
matrix(1:9, byrow = TRUE, nrow = 3 )
```

## Adding a column

In the previous exercise you calculated the vector that contained the worldwide box office receipt for each of the three Star Wars movies. However, this vector is not yet part of star_wars_matrix.

You can add a column or multiple columns to a matrix with the `r cbind()` function, which merges matrices and/or vectors together by column. For example:

```
big_matrix <- cbind(matrix1, matrix2, vector1 ...)
```
Instructions

Add worldwide_vector as a new column to the star_wars_matrix and assign the result to all_wars_matrix. Use the `r cbind()` function.

```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
region <- c("US", "non-US")
titles <- c("A New Hope", 
            "The Empire Strikes Back", 
            "Return of the Jedi")
               
star_wars_matrix <- matrix(box_office, 
                      nrow = 3, byrow = TRUE,
                      dimnames = list(titles, region))

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix,  worldwide_vector)
```

## Adding a row

Just like every action has a reaction, every `r cbind()` has an `r rbind()`. (We admit, we are pretty bad with metaphors.)

Your R workspace, where all variables you defined 'live' (check out what a workspace is), has already been initialized and contains two matrices:

```
* star_wars_matrix that we have used all along, with data on the original trilogy

* star_wars_matrix2, with similar data for the prequels trilogy.
```

Explore these matrices in the console if you want to have a closer look. If you want to check out the contents of the workspace, you can type `r ls()` in the console.


Instructions

Use `r rbind()` to paste together *star_wars_matrix*` and *star_wars_matrix2* , in this order. Assign the resulting matrix to *all_wars_matrix*.

```{r}
# star_wars_matrix and star_wars_matrix2 are available in your workspace
star_wars_matrix  

# star_wars_matrix 2
renglon <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")
columna <- c("US", "non-US")

star_wars_matrix2<- matrix(c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5),byrow = TRUE,nrow = 3,dimnames = list(renglon, columna))


# Combine both Star Wars trilogies in one matrix
(all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2))
```

## The total box office revenue for the entire saga

Just like *cbind()* has *rbind()*, *colSums()* has *rowSums()*. Your R workspace already contains the all_wars_matrix that you constructed in the previous exercise; type all_wars_matrix to have another look. Let's now calculate the total box office revenue for the entire saga.

Instructions

Calculate the total revenue for the US and the non-US region and assign total_revenue_vector. You can use the *colSums()* function.

Print out total_revenue_vector to have a look at the results.

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Total revenue for US and non-US
total_revenue_vector <- colSums(all_wars_matrix)
  
# Print out total_revenue_vector
total_revenue_vector
```

## Selection of matrix elements

Similar to vectors, you can use the square brackets [ ] to select one or multiple elements from a matrix. Whereas vectors have one dimension, matrices have two dimensions. You should therefore use a comma to separate the rows you want to select from the columns. For example:

* *my_matrix[1,2]* selects the element at the first row and second column.

* *my_matrix[1:3,2:4]* results in a matrix with the data on the rows 1, 2, 3 and columns 2, 3, 4.

If you want to select all elements of a row or a column, no number is needed before or after the comma, respectively:

* *my_matrix[,1]* selects all elements of the first column.

* *my_matrix[1,]* selects all elements of the first row.

Back to Star Wars with this newly acquired knowledge! As in the previous exercise, all_wars_matrix is already available in your workspace.

Instructions

* Select the non-US revenue for all movies (the entire second column of all_wars_matrix), store the result as non_us_all.

* Use mean() on non_us_all to calculate the average non-US revenue for all movies. Simply print out the result.

* This time, select the non-US revenue for the first two movies in all_wars_matrix. Store the result as non_us_some.

* Use mean() again to print out the average of the values in non_us_some.

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[ ,2]
  
# Average non-US revenue
mean(non_us_all)
#  the non-US revenue for first two movies
non_us_some <- all_wars_matrix[1:2,2]
  
# Average non-US revenue for first two movies
mean(non_us_some)

```

##A little arithmetic with matrices

Similar to what you have learned with vectors, the standard operators like +, -, /, *, etc. work in an element-wise way on matrices in R.

For example, **2 * my_matrix** multiplies each element of my_matrix by two.

As a newly-hired data analyst for Lucasfilm, it is your job to find out how many visitors went to each movie for each geographical area. You already have the total revenue figures in all_wars_matrix. Assume that the price of a ticket was 5 dollars. Simply dividing the box office numbers by this ticket price gives you the number of visitors.

Instructions

* Divide all_wars_matrix by 5, giving you the number of visitors in millions. Assign the resulting matrix to visitors.

* Print out visitors so you can have a look.
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Estimate the visitors
visitors <- all_wars_matrix / 5
  
# Print the estimate to the console
visitors
```

## A little arithmetic with matrices (2)

Just like **2 * my_matrix** multiplied every element of my_matrix by two, **my_matrix1 * my_matrix2** creates a matrix where each element is the product of the corresponding elements in my_matrix1 and my_matrix2.

After looking at the result of the previous exercise, big boss Lucas points out that the ticket prices went up over time. He asks to redo the analysis based on the prices you can find in ticket_prices_matrix (source: imagination).

Those who are familiar with matrices should note that this is not the standard matrix multiplication for which you should use **%*%** in R.

Instructions

* Divide all_wars_matrix by ticket_prices_matrix to get the estimated number of US and non-US visitors for the six movies. Assign the result to visitors.

* From the visitors matrix, select the entire first column, representing the number of visitors in the US. Store this selection as us_visitors.

* Calculate the average number of US visitors; print out the result.

```{r}

ticket_prices_matrix <- matrix(c(5, 5, 6, 6, 7, 7, 4, 4, 4.5, 4.5, 4.9, 4.9),byrow = TRUE,nrow = 6,dimnames = list(rownames(all_wars_matrix), columna))

# all_wars_matrix and ticket_prices_matrix are available in your workspace
all_wars_matrix
ticket_prices_matrix

# Estimated number of visitors
visitors <- all_wars_matrix / ticket_prices_matrix

# US visitors
us_visitors <- visitors[,1] 

# Average number of US visitors
mean(us_visitors)
```

